---
title: "Covid91 vaccine study Final2025"
author: "Jarrod Harris"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor)) 
males = filter(FinalData, sex =="M")
females = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid19) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

###Barplot

```{r}
dd2 <- males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took that vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= males)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of males that got covid-19 after taking the placebo and taking the vaccine. 
Among the guys, 26.20% got covid-19 after taking the vaccine, while the approximately 73.80% of guys that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in males and it is not recommended. Those who took the placebo were much less likely to contact covid-19
The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.

#Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 0. This is an awfully small number. seeing this number means that these results are not likely to occur by chance. The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option.

```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessmales = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessmales)
```


```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine for males. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we cn conclude that the vaccine is ineffective in protecting guys from getting covid-19. 
The odds ratio is approximately how effective the vaccine is. The odds ratio for the guys can be seen above as 70. The effectiveness confidence interval for the guys is 68.84 to 78.32.

## Females

### Barplot

```{r}
dd2 <- females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of females people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. In this plot, it appears that more people who were only given the placebo contracted covid-19. Only approximately 33% of the individuals who were given the vaccine were reported to have contracted covid-19. That means around 67% of the people that got covid-19 were the individuals that only received the placebo.The odds of getting covid-19 while taking the vaccine is 1:2. The gals that took the vaccine were half as likely to get covid-19 than those taking the placebo. There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not. In the second plot, the left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
=colPerc(table1)
```

These numerical results show specific percentages of females that got covid-19 after taking the placebo and taking the vaccine. 
The females that took the vaccine ans contracted covid-19 was 35%, just as approximated in the plots prior. The gals that took the placebo and got covid-19 was 65%. The vaccine is in fact effective in protecting females against the contraction of covid-19. 
The odds of getting covid-19 while taking the vaccine is 1:4 for females. The females that took the vaccine were half as likely to get covid-19 than those taking the placebo.

### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
femalesfish <- fisher.test(table1)
femalesfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is  small at 2e-08 . Seeing this number means that these results are likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.
The Fishers's Exact Test illustrates that the odds ratio is just under 2 at 1.9. Females who got the drug were about 95% chance to get covid-19 than women that only took the placebo.

```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemales = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemales)
```

```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 for females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is highly effective in protecting females from getting covid-19. The hallmark of an effective vaccine is 65% or better. This vaccine works great in protecting females from constructing covid-19. The odds ratio is approximately how effective the vaccine is. The odds ratio for the females can be seen above as 65. The effectiveness confidence interval for the females is from 1.49 to 2.32. 



## LGBTQ



## Druggies


# Overall Results and Conclusions